Reaction-diffusion spatial modeling of COVID-19 in Chicago

### 2021-07-27 draft/chicago-spatial-covid-draft.tex
Expand transmission rate discussion

@@

@@

Immediately we see there are infinitely many non-isolated fixed points of the form $(S,0)$.
Linearizing, we get
\begin{equation*}
\mathbf{A}_{(S,0)} = \begin{bmatrix} 0 & - \tilde{\beta} S \\ 0 & \beta S - \gamma \end{bmatrix}.
\end{equation*}
Thus, as seen in Figure \ref{fig:localdyn}, for $S < \frac{\gamma}{\tilde{\beta}}$ the nodes are stable, and for $S > \frac{\gamma}{\tilde{\beta}}$ they are unstable.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.25\textwidth]{local_dynamics}
	\caption{Stream plot for the local ODE dynamics. Note the transcritical bifurcation around $S_0 = 0.4$.}
	\label{fig:localdyn}
\end{figure}
We can then conclude that an epidemic will occur only if $S_0 = S(t=0) > \frac{\gamma}{\tilde{\beta}}$ \cite{Wang_SIR}.
Hence, a transcritical bifurcation occurs at $\Ro = \tilde{\beta} S_0 / \gamma$, which is known as the \textit{basic reproduction number}, and is used to estimate how the epidemic developed initially.

>>>>

The optimal parameters were determined for a scenario in which restrictive measures (quarantine, lockdown) were not enforced.
Therefore, we do not impose a time dependence on the transmission rate $\beta$.
The transmission rate represents the rate of contact of susceptible individuals with infectious individuals that leads to infection by the virus.
It may be expressed as the product of daily average contacts and the probability of infection given a contact \cite{Drossinos_2020}.
Thus, restrictive measures such as mobility restrictions, social-distancing, face-mask wearing, and gathering limitations are expected to decrease the transmission rate, an effect not accounted for in this model.

@@

The two top panels of Figure \ref{fig:ode_results} show how well we capture the data for the Chicago outbreak.
The model reproduces reasonable well the number of cases (top left panel) but not the number of deaths (top right panel).
The model fails to capture the characteristic S-shape in the logarithmic plot of the top right panel (deaths).

@@

Immediately we see there are infinitely many non-isolated fixed points of the form $(S,0)$.
Linearizing, we get
\begin{equation*}
\mathbf{A}_{(S,0)} = \begin{bmatrix} 0 & - \tilde{\beta} S \\ 0 & \beta S - \gamma \end{bmatrix}.
\end{equation*}
Thus, as seen in Figure \ref{fig:localdyn}, for $S < \frac{\gamma}{\tilde{\beta}}$ the nodes are stable, and for $S > \frac{\gamma}{\tilde{\beta}}$ they are unstable.
\begin{figure}[h]
	\centering
	\includegraphics[width=0.25\textwidth]{local_dynamics}
	\caption{Stream plot for the local ODE dynamics. Note the transcritical bifurcation around $S_0 = 0.4$.}
	\label{fig:localdyn}
\end{figure}
We can then conclude that an epidemic will occur only if $S_0 = S(t=0) > \frac{\gamma}{\tilde{\beta}}$ \cite{Wang_SIR}.
Hence, a transcritical bifurcation occurs at $\Ro = \tilde{\beta} S_0 / \gamma$, which is known as the \textit{basic reproduction number}, defined as the average number of secondary infections caused by a single infected individual \cite{Drossinos_2020}.
